---
title: "p8105_hw2_jl5549"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(tidyverse)
library(tidyr)
library(dplyr)
```

Problem 1
```{r} 
#import Mr Trash Wheel dataset
mr_trash_wheel = 
  read_excel( './data/HealthyHarborWaterWheelTotals2018-7-28.xlsx', 
                            sheet = 'Mr. Trash Wheel') %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  #way2 filter(is.na(dumpster) == 0)
  #way3 filter(!is.na(dumpster)) 
  mutate(sports_balls = round(sports_balls), 
         sports_balls = as.integer(sports_balls))
```

```{r}
#import percipitation datasets for 2017 and 2018
prcp_2017 = 
  read_excel('./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx', 
                            sheet = 4, skip = 1) %>% 
  janitor::clean_names() %>% 
  drop_na(total) %>% 
  mutate(year = '2017')
prcp_2018 = 
  read_excel('./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx', 
                            sheet = 3, skip = 1) %>% 
  janitor::clean_names() %>% 
  drop_na(total) %>% 
  mutate(year = '2018')

#combine two datasets
prcp = 
  rbind(prcp_2017, prcp_2018) %>% 
  mutate(month = month.name[as.integer(month)])

```

The number of the observations of first dataset is `r {nrow(mr_trash_wheel)}`. This dataset includes variables of dumpster number(dumpster), date of collection(year, month, date), amount of total litter(weight_tons) and litter type(weight_tons, sports_balls etc). 
The number of the observations of the second dataset is `r {nrow(prcp)}`. This dataset includes variables of precipitation amounts and associated dates(month, year). 
The total precipitation in 2018 is `r {sum(subset(prcp, year =='2018' & month != 'total')$total)}`. The median number of sports balls in a dumpster in 2017 is `r {median(subset(mr_trash_wheel, year = '2017')$sports_balls)}`.
The example is show as below.

```{r}
head(mr_trash_wheel)
head(prcp)
```


Problem 2

```{r}
#import pols-month dataset
plos_month = read_csv(file = './data/fivethirtyeight_datasets/pols-month.csv') %>% 
  janitor::clean_names() %>% 
  separate(mon, c('year','month','day'), sep = '-', convert = TRUE) %>% 
  
#create a president variable taking values gop and dem
#remove prez_dem, prez_gop and day
  mutate(month = month.name[as.integer(month)],
         president = replace(prez_gop, prez_gop == '1', 'gop'),
        president = replace(president, president == '0', 'dem')) %>% 
  select(-prez_gop, -prez_dem, -day) 

```

```{r} 
#import snp.csv dataset
#clean the data in snp.csv using a similar process to the above.
snp = read_csv(file = './data/fivethirtyeight_datasets/snp.csv') %>% 
  janitor::clean_names() %>% 
  separate(date, into = c('month', 'day', 'year'), sep = '/') %>% 
  select(year, month, everything()) %>% 
  mutate(month = month.name[as.integer(month)],
    year = as.integer(year)) %>% 
  arrange(year, month)

```

```{r}
#import snp.csv dataset
unemployment = read_csv(file = './data/fivethirtyeight_datasets/unemployment.csv') %>% 
  janitor::clean_names() %>% 
  pivot_longer(jan:dec, names_to = 'month', values_to = 'unemployment') %>% 
  mutate(month = as.character(factor(month, labels = month.name)),
         year = as.integer(year))

```

```{r}
#join datasets
merged_data = full_join(plos_month, snp, by = "year") %>% 
  full_join(unemployment, by = 'year')

```
 
The first dataset plos_month contains datas from 1947 to 2015. It contains `r {nrow(plos_month)}`  rows and `r {ncol(plos_month)}` variables related to the number of national politicians who are democratic or republican at any given time. The president variable indicate whether the president was republican or democratic on the associated date. The variable gov_gop is the number of republican governors on the associated date. The variable sen_gop is the number of republican senators on the associated date. The variable rep_gop is the number of republican representatives on the associated date. The variable gov_dem is the number of democratic governors on the associated date. The variable sen_dem is the number of democratic senators on the associated date. The variable rep_dem is the number of democratic representatives on the associated date.

The second dataset snp contains datas from 1950 to 2014. It contains `r {nrow(snp)}` rows and `r {ncol(snp)}` variables related to representative measure of stock market as a whole. The date variable is the date of the observation. The close variable is the closing values of the S&P stock index on the associated date. 

The second dataset unemployment contains datas from 1948 to 2015. It contains `r {nrow(unemployment)}` rows and `r {ncol(unemployment)}` variables to indicate the employment rate on a specific date. 
The last dataset is a merged on by the three above which shows comprehensive information related to election of the associated year. It contains `r {nrow(merged_data)}` rows and `r {ncol(merged_data)}` variables where variable year is the year of the measurements on that row.

Problem 3
```{r}
#import dataset
#address the issues of changing names of a categorical predictor and the case structure of string variables
#remove duplicate rows
baby_names = read_csv(file = './data//Popular_Baby_Names.csv') %>% 
  janitor::clean_names() %>% 
  mutate(childs_first_name = str_to_lower(childs_first_name),
         first_letter = str_to_upper(substr(childs_first_name, 1, 1)),
         childs_first_name = str_c(first_letter, substr(childs_first_name,2,nchar(childs_first_name))),
         ethnicity = recode(ethnicity, 'ASIAN AND PACI' = 'ASIAN AND PACIFIC ISLANDER', 
                            'BLACK NON HISP' = 'BLACK NON HISPANIC', 
                            'WHITE NON HISP' = 'WHITE NON HISPANIC')
          ) %>%
  select(-first_letter) %>% 
  distinct()

```

```{r}
#the rank of “Olivia” over time
olivia = subset(baby_names, childs_first_name == 'Olivia', 
                select =c(year_of_birth, ethnicity, rank)) %>% 
  pivot_wider(names_from = 'year_of_birth', values_from = 'rank')

#the most popular male children name over time
male_popular = subset(baby_names, rank == '1' & gender == 'MALE',
                      select =c(year_of_birth, ethnicity, childs_first_name)) %>% 
  pivot_wider(names_from = 'year_of_birth', values_from = 'childs_first_name')

```

```{r}
# scatter plot showing the number of children with a name against the rank in popularity of that name
non_hispanic = subset(baby_names, ethnicity=='WHITE NON HISPANIC' & gender == 'MALE' & year_of_birth == '2016', 
       select = c(childs_first_name, rank, count))
ggplot(non_hispanic, aes(x = rank, y = count))+
  geom_point(color = 'pink')

```

